Epistemic Logics For Knowledge Bases

### 2010-12-15 introduction.tex
Refactored paper so all sections are seperate documents

\documentclass{article}
\usepackage{geometry,amsmath,amssymb,wasysym,enumerate,bbm,color}
\geometry{letterpaper}

\usepackage{stmaryrd}

%%%%%%%%%% Start TeXmacs macros
\newcommand{\assign}{:=}
\newcommand{\nin}{\not\in}
\newcommand{\tmem}[1]{{\em #1\/}}
\newcommand{\tmop}[1]{\ensuremath{\operatorname{#1}}}
\newcommand{\tmstrong}[1]{\textbf{#1}}
\newcommand{\tmtextbf}[1]{{\bfseries{#1}}}
\newcommand{\tmtextit}[1]{{\itshape{#1}}}
\newcommand{\tmtextsf}[1]{{\sffamily{#1}}}
\newcommand{\tmtextup}[1]{{\upshape{#1}}}
\newenvironment{descriptioncompact}{\begin{description} }{\end{description}}
\newenvironment{descriptiondash}{\begin{description} }{\end{description}}
\newenvironment{enumeratenumeric}{\begin{enumerate}[1.] }{\end{enumerate}}
\newenvironment{enumerateroman}{\begin{enumerate}[i.] }{\end{enumerate}}
\newenvironment{itemizedot}{\begin{itemize} \renewcommand{\labelitemi}{$\bullet$}\renewcommand{\labelitemii}{$\bullet$}\renewcommand{\labelitemiii}{$\bullet$}\renewcommand{\labelitemiv}{$\bullet$}}{\end{itemize}}
\newenvironment{proof}{\noindent\textbf{Proof\ }}{\hspace*{\fill}$\Box$\medskip}
\definecolor{grey}{rgb}{0.75,0.75,0.75}
\definecolor{orange}{rgb}{1.0,0.5,0.5}
\definecolor{brown}{rgb}{0.5,0.25,0.0}
\definecolor{pink}{rgb}{1.0,0.5,0.5}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
%%%%%%%%%% End TeXmacs macros

\begin{document}

\title{Epistemic Logics For Knowledge Bases}\author{Matthew P.
Wampler-Doty}\maketitle

\begin{abstract}
  In {\cite{van_benthem_reflectionsepistemic_1991}}, Johan van Benthem
  proposed a research program for logic of \tmtextit{explicit knowledge}. \
  Two efforts have since emerged with proposals for logic of explicit
  knowledge: the \tmtextit{Dynamics of Awareness}
  {\cite{van_benthem_inference_2009}} and \tmtextit{Justification Logic}
  {\cite{artemov_introducing_2005}}.  The purpose of this paper is to
  repurpose various epistemic logics from these previous efforts for reasoning
  about logical closures over knowledge bases.  The results in this paper are
  based on previous work carried out in
  {\cite{wampler-doty_evidentialist_2010}}.
\end{abstract}

\section{Introduction\label{intro}}

The modern tradition in epistemic logic is to assume knowledge modalities
conform to the $S 5$ axiom schema.  As noted in
{\cite{halpern_set-theoretic_1999,rubinstein_modeling_1998}}, the semantics of
$S 5$ knowledge correspond exactly to partitioning a collection of situations
into \tmtextit{information sets}, which is the traditional approach in game
theory and decision theory.  While it is not commonly acknowledged in
epistemic logic, economists and philosophers accept that traditional decision
theory is externalist and behaviourist in nature
(FIXME)
\footnote{An early essay by
Amartya Sen on the philosophical foundations of traditional decision theory
makes the behaviorist reading of decision theory clear
\cite{sen_behaviour_1973}.  Kaushik Basu also discusses the behaviorist
nature of decision theory {\cite[pgs. 53--54]{basu_revealed_1980}}. 
Finally, Donald Davidson appeals to decision theory to motivate an externalist
epistemology in \cite{davidson_could_1995}.}.

In {\cite{van_benthem_reflectionsepistemic_1991}}, Johan van Benthem proposed
a research program to find logics for explicit knowledge, providing the first
suggestion of an internalist perspective on epistemic logic.  Subsequently,
Sergei Artemov and Elena Nogina proposed a logic of explicit justification,
which has come to be known as \tmtextit{Justification Logic} (JL)
{\cite{artemov_introducing_2005}}, based on Artemov's \tmtextit{Logic of
Proofs} {\cite{artemov_logic_1994}}.  While the original semantics of JL was
interpretability into Peano Arithmetic, Melvin Fitting proposed Kripke
semantics for JL in {\cite{fitting_logic_2005}}.  Recently, Fernando
Vel\'azquez-Quesada and Johan van Benthem have developed a simpler framework
for explicit epistemics in {\cite{van_benthem_inference_2009}}, entitled the
\tmtextit{Dynamics of Awareness}.  This work was based on Joseph Halpern and
Ronald Fagin's original \tmtextit{Awareness Logic} {\cite{fagin_belief_1987}}.

@@

We adopt a arguments-on-paper-thesis perspective on epistemic logic in this
paper.  We consider a \tmtextit{good argument} to be a logical derivation
from propositions present in a knowledge base.  In this fashion, ou.  \
Special attention will be given to \tmtextit{sound derivations}, which will be
thought of as a form of knowledge.  In Awareness Logic, we will interpret
awareness of a formula as membership in a knowledge base.  We prove
completeness for basic awareness logic and a hybrid logic extension.  Logics
of multiple knowledge bases are also presented: a simplified form of
Justification Logic, a logic with neighborhood semantics, an a logic with
modalities for quantifying over knowledge bases.  We conclude with an
application to naturalized epistemology found in the psychological literature.

\section{Basic Explicit Logics of Knowledge Bases}

\subsection{Awareness Logic\label{awarenesslogic}}

In this section, we give the explicit semantics to basic awareness logic,
originally presented in {\cite{fagin_belief_1987}}, with an additional axiom.
\ One difference in our presentation is the inclusion of a novel letter and a
superficial change in notation.  This work is inspired by the developments in
van Benthem and Vel\'azquez-Quesada {\cite{van_benthem_inference_2009}}.

\begin{definition}
  Let $\Phi$ be a set of letters and define the language $\mathcal{L}_A
  (\Phi)$ as:
  \[ \phi \hspace{1em} : : = \hspace{1em} p \in \Phi \hspace{1em} |
     \hspace{1em} \circlearrowleft \hspace{1em} | \hspace{1em} \bot
     \hspace{1em} | \hspace{1em} \phi \rightarrow \psi \hspace{1em} |
     \hspace{1em} \Box \phi \hspace{1em} | \hspace{1em} A : \phi \hspace{1em}
  \]
\end{definition}

We define the other connectives and modal operators as usual for classical
modal logics.

\begin{definition}
  \label{awarenessmodels}An {\tmstrong{awareness model}} $\mathbbm{M}= \langle
  W, V, R, \mathcal{A} \rangle$ is a Kripke model with a valuation $V : \Phi
  \cup \{\circlearrowleft\} \rightarrow 2^W$ and an awareness function
  $\mathcal{A} : W \rightarrow 2^{\mathcal{L}_A (\Phi)}$.  We will usually
  denote $\mathcal{A} (w)$ as $\mathcal{A}_w$.

  Define the semantic entailment relation $\vDash$ where atomic propositions,
  logical connectives and modality are interpreted as usual.  We employ the
  following novel definition:
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash A : \phi & \text{iff} & \phi \in \mathcal{A_w}
  \end{eqnarray*}
\end{definition}

The following definition distinguishes us from previously investigations of
awareness logic:

\begin{definition}
  Define the ``\tmtextbf{theory of a model} $\mathbbm{M}$'' as:
  \[ \tmop{Th} (\mathbbm{M}) \assign \{\phi \hspace{1em} | \hspace{1em}
     \mathbbm{M}, v \vDash \phi \text{ for all } v \in W\} \]
  We define the following properties that a model $\mathbbm{M}$ may posses:

  \begin{descriptiondash}
    \item[CSQ] $\mathbbm{M}, w \vDash \Box \phi \text{ iff } \tmop{Th}
    (\mathbbm{M}) \cup \mathcal{A}_w \vdash \phi$

    \item[SND] $\mathbbm{M}, w \vDash \circlearrowleft \text{ iff }
    \mathbbm{M}, w \vDash \phi$ for all $\phi \in \mathcal{A}_w$
  \end{descriptiondash}

  Here $\vdash$ is any logical consequence relation which is sound for
  $\vDash$, making true \tmtextbf{modus ponens} and \tmtextbf{reflection} (ie,
  $\Gamma \vdash \phi$ if $\phi \in \Gamma$)
\end{definition}

Neither \tmtextbf{CSQ} nor \tmtextbf{SND} correspond to simple modally
definable properties.

Intuitively, the \tmtextbf{CSQ} asserts that an agent believes a formula
$\phi$ if and only if it follows logically from their knowledge base and their
background knowledge, represented by $\tmop{Th} (\mathbbm{M})$. \
\tmtextbf{CSQ} is Vincent Hendrik's principle of logical omniscience mentioned
in {\S}\ref{intro}.

\tmtextbf{SND} asserts that $\circlearrowleft$ corresponds to the agent's
knowledge base being sound. A sound knowledge base will only render true
conclusions when used in deductions.  If one considers ``the existance of a
sound deduction'' sufficient for a kind of knowledge, then $\circlearrowleft$
is suitable for investigating this notion of epistemology.  That is, if
$\mathbbm{M}, w \vDash \circlearrowleft \wedge \Box \phi$, then we may
naturally say that the agent (implicitly) {\tmem{knows}} $\phi$ on the basis
of $\mathcal{A}_w$, rather than merely {\tmem{believing}} $\phi$ on the basis
of $\mathcal{A}_w$.

The logic of models making true \tmtextbf{CSQ} and \tmtextbf{SND} is provided
in Table \ref{logic1}.

\begin{table}[h]
  \begin{tabular}{l}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$\\
    $\vdash \Box(\phi \rightarrow \psi) \rightarrow \Box \phi \rightarrow \Box
    \psi$\\
    $\vdash A : \phi \rightarrow \Box \phi$\\
    $\vdash \circlearrowleft \rightarrow \Box \phi \rightarrow \phi$\\
    \\
    \begin{tabular}{lll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ & {\hspace{6em}} & $\frac{\vdash \phi}{\vdash \Box \phi}$
    \end{tabular}
  \end{tabular}
  \caption{\label{logic1}Awareness Logic for \tmtextbf{CSQ} and
  \tmtextbf{SND}}
\end{table}

\begin{theorem}
  \label{completeness1}Assuming an infinite store of proposition letters
  $\Phi$, the awareness logic presented is sound and weakly complete for
  awareness models making true \tmtextbf{CSQ} and \tmtextbf{SND}.
\end{theorem}

\begin{proof}
  Soundness is straightforward, so we will only address completeness.  Assume
  $\nvdash \psi$.  Consider the finite canonical model $\mathbbm{M}= \langle
  W, V, R, \mathcal{A} \rangle$ formed of maximally consistent sets of
  subformulae of $\phi$ (closed under single negations{\footnote{A discussion
  of this property may be found in {\cite[pg.
  243]{blackburn_modal_2001}}.}}), as per the finitary modal completeness
  proofs found in {\cite[chapter 4.8]{blackburn_modal_2001}} and
  {\cite[chapter 5]{boolos_logic_1995}}.  Typical presentations do not
  include awareness, however it is defined intuitively in this case:
  \[ \phi \in \mathcal{A}_w \Longleftrightarrow A : \phi \in w \]
  Since $\psi$ is a subformula of itself, we have $\mathbbm{M}, w \nvDash
  \psi$ for some world $w$ by a finitary {\tmem{Lindenbaum Lemma}} and
  {\tmem{Truth Theorem}}.  Moreover, it is straightforward to verify that
  $\mathbbm{M}$ makes true the following properties:
  \begin{enumeratenumeric}
    \item $W$ and $\mathcal{A}_v$ are finite, and $\psi \in \mathcal{A}_v$
    only if $\psi$ is a (possibly negated) subformula of $\phi$

    \item if $\mathbbm{M}, v \vDash A : \psi$ then $\mathbbm{M}, v \vDash \Box
    \psi$

    \item if $\mathbbm{M}, v \vDash \circlearrowleft$ then $v R v$
  \end{enumeratenumeric}

  We next produce a bisimular{\footnote{The notion of bisimulation is
  discussed at length in {\cite[chapter 2.2]{blackburn_modal_2001}}.  We
  note that the basic definition and all of the usual theorems (modal
  equivalence, Henessy Milner, etc.) may be generalized to awareness models in
  a straightforward fashion. }} model $\mathbbm{M}'$ which makes true (1), (2)
  and a stronger from of (3):

  {\hspace*{\fill}}\begin{tabular}{ll}
    \tmtextbf{3$'$.} & $\mathbbm{M}', v \vDash \circlearrowleft$ iff $v R v$
  \end{tabular}{\hspace*{\fill}}

  To this end define $\mathbbm{M}' \assign \langle W', V', R', \mathcal{A}'
  \rangle$ such that{\footnote{Throughout this article, we use $l$ and $r$ to
  denote the two canonical injections associated with the coproduct $W \uplus
  W$. We will use $v_l$ and $v_r$ as the shorthand for $l (v)$ and $r (v)$
  respectively.}}: {\hspace*{\fill}}
  \begin{itemizedot}
    \item $W' \assign W \uplus W$

    \item $V' (p) \assign \{v_l, v_r \hspace{1em} | \hspace{1em} v \in V
    (p)\}$

    \item $R' \assign \{(v_l, u_r), (v_r, u_l) \hspace{1em} | \hspace{1em} v R
    u\} \cup \{(v_l, v_l), (v_r, v_r) \hspace{1em} | \hspace{1em} \mathbbm{M},
    v \vDash \circlearrowleft\}$

    \item $\mathcal{A}' (v_l) \assign \mathcal{A}' (v_r) \assign \mathcal{A}
    (v)$
  \end{itemizedot}

  It is straightforward to verify that $\mathbbm{M}'$ makes true the desired
  properties.  Let $Z \assign \{(v, v_l), (v, v_r) \hspace{1em} |
  \hspace{1em} v \in W\}$.  then $Z$ is a bisimulation between $\mathbbm{M}$
  and $\mathbbm{M}'$.  Therefore we know there is some $w \in W$ such that
  $\mathbbm{M}', w_l \nvDash \psi$ and $\mathbbm{M}', w_r \nvDash \psi$.

  Finally we construct a model $\mathbbm{M}''$ which agrees with
  $\mathbbm{M}'$ for all subformulae of $\psi$, which makes true
  \tmtextbf{CSQ} and \tmtextbf{SND}.  Let $\Lambda$ be the set of proposition
  letters occuring in $\psi$.  We will make use members of $\Phi \backslash
  \Lambda$ as {\tmem{nominals}}, as per the tradition in hybrid logic.  This
  is possible since both $W'$ and $\Lambda$ are finite and $\Phi$ is infinite
  by hypothesis.  Let $\iota : W \uplus W \hookrightarrow \Phi \backslash
  \Lambda$ be an injection assigning a fresh nominal associated with each
  world in $\mathbbm{M}'$.  Now define $\mathbbm{M}'' \assign \langle W'',
  V'', R'', \mathcal{A}'' \rangle$, where:

  \begin{center}
    \begin{tabular}{lllll}
      $\bullet$ & $W'' \assign W'$ &  & $\bullet$ & $V'' (p) \assign \left\{
      \begin{array}{ll}
        V' (p) & p \in \Lambda\\
        \{v\} & p = \iota (v)\\
        \varnothing & o / w
      \end{array} \right.$\\
      $\bullet$ & $R'' \assign R'$ & {\hspace{3em}} & $\bullet$ &
      $\mathcal{A}''_v \assign \{\phi \hspace{1em} | \phi \in \mathcal{A}'_v
      \} \cup \{\neg \iota (u) \hspace{1em} | \hspace{1em} \neg v R' u\}$
    \end{tabular}
  \end{center}

  By induction we have that for every subformula $\phi$ of $\psi$ that
  $\mathbbm{M}', v \vDash \phi$ if and only if $\mathbbm{M}'', v \vDash \phi$,
  hence $\mathbbm{M}'', w_l \nvDash \psi$.

  All that is left to show is that $\mathbbm{M}''$ makes true \tmtextbf{CSQ}
  and \tmtextbf{SND}.  $\mathbbm{M}''$ has three further properties:
  \begin{enumerateroman}
    \item $\mathcal{A}''_v$ is finite for all worlds $v$

    \item  $v R' u$ if and only if $\mathbbm{M}'', u \vDash \bigwedge
    \mathcal{A}'' (v)$

    \item The logic presented in Table \ref{logic1} is sound for
    $\mathbbm{M}''$
  \end{enumerateroman}
  From (ii) and the fact that $\mathbbm{M}'$ makes true (3$'$), we have
  \tmtextbf{SND} for $\mathbbm{M}''$.

  All that is left is to demonstrate \tmtextbf{CSQ} for some sound logic.  We
  will use the logic in Table \ref{logic1} itself.  From (i), (ii), (iii),
  and the \tmtextit{deduction theorem} for modal logic, we have the following
  line of reasoning:

  \begin{align*}
    Th({\mathbbm{M}}''){\cup}\mathcal{A}''_v{\vdash}{\phi} &
    {\Longleftrightarrow}Th({\mathbbm{M}}''){\vdash}\bigwedge\mathcal{A}''_v{\rightarrow}{\phi}\\
    &
    {\Longleftrightarrow}\bigwedge\mathcal{A}''_v{\rightarrow}{\phi}{\in}Th({\mathbbm{M}}'')\\
    &
    {\Longleftrightarrow}{\mathbbm{M}}'',u{\vDash}\bigwedge\mathcal{A}''_v{\rightarrow}{\phi}\text{
    for all $u \in W''$}\\
    & {\Longleftrightarrow}\text{ for all $u \in W''$, if
    }{\mathbbm{M}}'',u{\vDash}\mathcal{A}''_v\text{ then
    }{\mathbbm{M}}'',u{\vDash}{\phi}\\
    & {\Longleftrightarrow}\text{ for all $u \in W''$, if }v R 'u \text{ then
    }{\mathbbm{M}}'',u{\vDash}{\phi}\\
    & {\Longleftrightarrow}{\mathbbm{M}}'',v{\models}\Box{\phi}
  \end{align*}
\end{proof}

\subsection{Hybrid Awareness Logic}

The method of the above completeness proof makes implicit use of concepts from
hybrid logic{\footnote{Hybrid logic was first presented in
{\cite{prior_revised_1969}} and later formally developed in
{\cite{bull_approach_1970}}.}}.  In this section we adapt hybrid logic and
show that it is a complete for reasoning about knowledge bases.  We note that
because of the increased expressive power of hybrid logic, our completeness
theorem is simpler.

\begin{definition}
  Let $\Phi$ be a set of letters and $\Psi$ a set of nominals, and define the
  language $\mathcal{L}_H (\Phi, \Psi)$ as:
  \[ \phi \hspace{1em} : : = \hspace{1em} p \in \Phi \hspace{1em} |
     \hspace{1em} i \in \Psi \hspace{1em} | \hspace{1em} \bot \hspace{1em} |
     \hspace{1em} \phi \rightarrow \psi \hspace{1em} | \hspace{1em} \Box \phi
     \hspace{1em} | \hspace{1em} A : \phi \hspace{1em} | \hspace{1em} \forall
     \phi \]
\end{definition}

Our approach in hybrid logic is to employ a universal modality along with
nominals.  This framework presents a logic where the agent may reason about
various labeled scenarios.  Scenarios may be multiply labeled or not labeled
at all.  From these intuitions we have the following definition:

\begin{definition}
  \label{hybridsemantics}Let a \tmtextbf{hybrid model} $\mathbbm{M}= \langle
  W, V, R, \mathcal{A}, \ell \rangle$ be an awareness model as in Definition
  \ref{awarenessmodels}, along with a partial labeling function $\ell : \Psi
  \nrightarrow W$

  The semantics for $\vDash$ are the same as in Definition
  \ref{awarenessmodels}, only we have
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash \forall \phi & \text{iff} & \forall v \in
    W.\mathbbm{M}, v \vDash \phi\\
    \mathbbm{M}, w \vDash i & \text{iff} & \ell (i) = w
  \end{eqnarray*}
\end{definition}

The other connectives and operators are defined as usual, although we employ a
few special shorthands:

\begin{definition}
  We employ the following special abbreviations:
  \begin{eqnarray}
    \text{\tmtextup{@}}_i \phi \assign \forall (i \rightarrow \phi)
    \hspace{2em} & \exists \phi \assign \neg \forall \neg \phi \hspace{3em} &
    \circlearrowleft_i \assign \diamondsuit i \nonumber
  \end{eqnarray}
\end{definition}

This gives rise to a validity reflecting one of the axioms we saw in
{\S}\ref{awarenesslogic}:
\[ \vDash \circlearrowleft_i \rightarrow \Box \phi \rightarrow
   \text{\tmtextup{@}}_i \phi \]
This can be read as ``If the agent's knowledge base is sound at world $i$,
then if they can deduce something from their knowledge base, what they deduce
must be true at world $i$.'.  In a way, this may be considered as
{\tmem{relativising}} knowledge particular named situations.

The semantics in Definition \ref{hybridsemantics} obviates the \tmtextbf{SND}
principle we previously presented, since there is no explicit
$\circlearrowleft$ symbol in this setting. \tmtextbf{CSQ} still makes sense
without modification, however.  The following gives a logic for hybrid models
making true \tmtextbf{CSQ}:

\begin{table}[h]
  \begin{tabular}{ll}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$ & $\vdash \forall (\phi
    \rightarrow \psi) \rightarrow \forall \phi \rightarrow \forall \psi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$ & $\vdash \forall
    \phi \rightarrow \phi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$ & $\vdash \forall \phi \rightarrow
    \forall \forall \phi$\\
    $\vdash \Box(\phi \rightarrow \psi) \rightarrow \Box \phi \rightarrow \Box
    \psi$ & $\vdash \exists \phi \rightarrow \forall \exists \phi$\\
    $\vdash A : \phi \rightarrow \Box \phi$ & $\vdash \forall \phi \rightarrow
    \Box \phi$\\
    & $\vdash i \rightarrow \phi \rightarrow \text{\tmtextup{@}}_i \phi$\\
    & \\
    \begin{tabular}{llll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ & {\hspace{6em}}$\frac{\vdash \phi}{\vdash \Box \phi}$ &
      {\hspace{6em}} & $\frac{\vdash \phi}{\vdash \forall \phi}$
    \end{tabular} & 
  \end{tabular}
  \caption{\label{logic2}Hybrid Awareness Logic for \tmtextbf{CSQ}}
\end{table}

\begin{theorem}
  \label{completeness2}Assuming an infinite store of nominals $\Psi$, the
  hybrid awareness logic presented is sound and weakly complete for all hybrid
  models making true \tmtextbf{CSQ}.
\end{theorem}

\begin{proof}
  As before, soundness is trivial.  Completeness is based on a modification
  of techniques found in {\cite[chapter 5]{boolos_logic_1995}}.  Assume
  that $\nvdash \psi$ and let $\Sigma_0$ be the set of subformulae of $\psi$,
  and let $\Upsilon$ be the set of nominals occuring in $\psi$.  Define:

  \begin{align*}
    {\Sigma}_1 &
    {\assign}{\Sigma}_0{\cup}\{{\forall}{\phi}{\hspace{1em}}|{\hspace{1em}}\Box{\phi}{\in}{\Sigma}_0\}\\
    {\Sigma}_2 &
    {\assign}{\Sigma}_1{\cup}\{\text{@}_i{\phi}{\hspace{1em}},i{\rightarrow}{\phi}|{\hspace{1em}}{\phi}{\in}{\Sigma}_1{\hspace{1em}}\&{\hspace{1em}}i{\in}{\Upsilon}\}\\
    {\Sigma}_3 &
    {\assign}{\Sigma}_2{\cup}\{{\neg}{\phi}{\hspace{1em}}|{\hspace{1em}}{\phi}{\in}{\Sigma}_2\}
  \end{align*}

  It is easy to see that $\Sigma_3$ is finite and closed under subformulae and
  single negations. Let $\mathbbm{M}= \langle W, V, R_{\Box}, \sim_{\forall},
  \mathcal{A} \rangle$ be the finite cannonical model formed of maximally
  consistent subsets of $\Sigma_3$, where everything is defined as usual,
  except $\sim_{\forall}$, which obeys the following specification:

  \begin{align*}
    w {\sim}_{{\forall}}v & {\Longleftrightarrow}(({\forall}{\phi}){\in}w
    {\Longleftrightarrow}({\forall}{\phi}){\in}v)
  \end{align*}

  This model makes true the following properties:
  \begin{enumerateroman}
    \item $\sim_{\forall}$ is an equivalence relation

    \item $R_{\Box} \subseteq \sim_{\forall}$

    \item In each equivalence class specified by $\sim_{\forall}$, there is at
    most one world making true $i$ for all $i \in \Upsilon$
  \end{enumerateroman}
  As in our previous construction we have that there is some world $w$ where
  $\mathbbm{M}, w \nvDash \psi$.  Let $\mathbbm{M}^1 = \langle W', V',
  R_{\Box}', \sim_{\forall}', \mathcal{A}' \rangle$ be the submodel generated
  by $\{w\}$; we have that $\mathbbm{M}', w \nvDash \psi$ (see {\cite[chapter
  2.1]{blackburn_modal_2001}} for details on generated submodels).  In this
  model $\forall$ is a universal modality and either $\left\llbracket i
  \right\rrbracket^{\mathbbm{M}'} = \varnothing$ or $\left\llbracket i
  \right\rrbracket^{\mathbbm{M}'} =\{v\}${\footnote{As per the usual
  convention, here $\left\llbracket \phi \right\rrbracket^{\mathbbm{M}}$
  denotes $\{w \in W \hspace{1em} | \hspace{1em} \mathbbm{M}, w \vDash
  \phi\}$.  We will drop $\mathbbm{M}$ where it is unambiguous.}}. Since the
  store $\Psi$ of nominals is infinite and $\Upsilon$ is finite, we have that
  $\Psi \backslash \Upsilon$ is infinite, so there is some injection $\iota :
  W' \hookrightarrow \Psi \backslash \Upsilon$ which assigns a fresh nominal
  to every world.  Let $\mathbbm{M}'' = \langle W'', V'', R'', \mathcal{A}'',
  \ell \rangle$, where:

  \begin{center}
    \begin{tabular}{lllll}
      $\bullet$ & $W'' \assign W'$ &  & $\bullet$ & $V'' (p) \assign V' (p)$\\
      $\bullet$ & $R'' \assign R'_{\Box}$ & {\hspace{3em}} & $\bullet$ &
      $\mathcal{A}'' (v) \assign \{\phi \hspace{1em} | \hspace{1em} \phi \in
      \mathcal{A}' (v)\} \cup \{\neg \iota (u) \hspace{1em} | \hspace{1em}
      \neg v R' u\}$\\
      \begin{tabular}{ll}
        $\bullet$ & $\ell (i) \assign \left\{ \begin{array}{ll}
          w & i \in \Upsilon \hspace{1em} \& \hspace{1em} \mathbbm{M}, w
          \vDash i\\
          \iota^{- 1} (i) & i \in \iota [W]\\
          \text{undefined} & \text{o/w}
        \end{array} \right.$
      \end{tabular} & $R'' \assign R'_{\Box}$ &  &  & 
    \end{tabular}
  \end{center}

  The the worlds is $\mathbbm{M}'$ and $\mathbbm{M}''$ agree on all
  subformulae of $\psi$, hence $\mathbbm{M}'', w \nvDash \psi$. The same
  reasoning as in the proof of Theorem \ref{completeness1} establishes and
  that the logic in Table \ref{logic2} itself is sound for $\mathbbm{M}''$ and
  makes true \tmtextbf{CSQ}, establishing completeness.
\end{proof}

\section{Logics of Multiple Knowledge Bases}

So far, we have only dealt with systems that reason over a single knowledge
base.  In this section, we present various logics for reasoning over multiple
knowledge bases.

\subsection{Simple Justification Logic}

Justification Logic (JL) was originally developed by Artemov as the
\tmtextit{Logic of Proofs} (LP) {\cite{artemov_logic_1994}}.  Its original
purpose was to provide a framework for reasoning about explicit provability in
Peano Arithmetic.  The first introduction of Justification Logic can be found
in {\cite{artemov_introducing_2005}}, where Artemov and Nogina propose LP as a
logic for reasoning about evidence.  In {\cite{fitting_logic_2005}}, Melvin
Fitting provided JL with Kripke model based semantics.

The principle innovation of LP/JL is to extend awareness logic, such that
awareness operations are now \tmtextit{proof terms}.  Informally, we say that
``a proof term $X$ witnesses proposition $\phi$'', denoted ``$X : \phi$'',
whenever $X$ represents a proof of $\phi$.  Proofs may have
{\tmstrong{multiple-conclusions}} in this system, so it is possible that $X :
\phi$ and $X : \psi$ can be true where $\phi \neq \psi$.  In Fitting's Kripke
semantics, $X : \phi$ means that $\phi$ is in awareness set corresponding to
$X$.  Proof terms are thought be operating in a \tmtextit{multi-conclusion}
proof system, so the same proof term may witness many different propositions.

The logics LP/JL include operators over proofs terms, so that new proof terms
may be assembled.  One operation of particular interest to us is
{\tmem{choice}}, denoted $\oplus$.  The expression ``$X \oplus Y : \phi$''
denotes that either $X$ or $Y$ are proofs witnessing $\phi$.  There are other
operations which correspond to \tmtextit{modus ponens} and
\tmtextit{proof-theoretic reflection}, however we do not consider these
operations here.

In this section we consider a simplified form of JL suitable for reasoning
over multiple knowledge bases, which we call \tmtextit{simple justification
logic} (SJL). However, instead of thinking of terms as representing proofs, we
use them to represent knowledge bases. This is in the spirit of JL as a logic
of evidence: we consider each knowledge base to be a corpus of evidence. \
Each term ``names'' a different knowledge base at a particular world.  Terms
may not refer to the same knowledge base at different worlds, just as in how
in awareness logic the agent need not be aware of the same formulae at
different worlds.  Finally, we will make use of JL's {\tmem{choice}} operator
as a mechanism for forming the union of knowledge bases, creating new ones.

\begin{definition}
  Let $\Pi$ be a set of primitive terms.  Define $\tau (\Pi)$ with the
  following grammar:
  \[ X \hspace{1em} : : = \hspace{1em} t \in \Pi \hspace{1em} | \hspace{1em} X
     \oplus Y \]

  Let $\Phi$ be a set of letters and $\Pi$ a set of primitive terms, and
  define the language $\mathcal{L}_{\tmop{SJL}} (\Phi, \Pi)$ as:
  \[ \phi \hspace{1em} : : = \hspace{1em} p \in \Phi \hspace{1em} |
     \hspace{1em} \circlearrowleft_X \hspace{1em} | \hspace{1em} \bot
     \hspace{1em} | \hspace{1em} \phi \rightarrow \psi \hspace{1em} |
     \hspace{1em} \Box_X \phi \hspace{1em} | \hspace{1em} X : \phi \]
  where $X \in \tau (\Pi)$
\end{definition}

\begin{definition}
  \label{justmodels}A {\tmstrong{simple justification model}} $\mathbbm{M}=
  \langle W, V, R, \mathcal{A} \rangle$ is a Kripke model with a valuation $V
  : \Phi \cup \{\circlearrowleft_X \hspace{1em} | \hspace{1em} X \in \tau
  (\Pi)\} \rightarrow 2^W$, an indexed relation $R : \tau (\Pi) \rightarrow
  2^{W \times W}$, along with a modified awareness function $\mathcal{A} : W
  \times \tau (\Pi) \rightarrow 2^{\mathcal{L}_{\tmop{SJL}} (\Phi, \Pi)}$. \
  In practice we will denote $\mathcal{A} (w, X)$ by a curried shorthand,
  namely $\mathcal{A}_w (X)$.

  The semantics for $\vDash$ have the following modifications:
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash \Box_X \phi & \text{iff} & \text{for all $v \in W$
    where $w R_X v$ we have $\mathbbm{M}, v \vDash v$}\\
    \mathbbm{M}, w \vDash X : \phi & \text{iff} & \phi \in \mathcal{A}_w (X)
  \end{eqnarray*}
\end{definition}

\begin{definition}
  The following defines properties a simple justification model may make true:

>>>>

The modern tradition in epistemic logic is to assume knowledge modalities
conform to the $S 5$ axiom schema.  As noted in
{\cite{halpern_set-theoretic_1999,rubinstein_modeling_1998}}, the semantics of
$S 5$ knowledge correspond exactly to partitioning a collection of situations
into \tmtextit{information sets}, which is the traditional approach in game
theory and decision theory.  While it is not commonly acknowledged in
epistemic logic, economists and philosophers accept that traditional decision
theory is externalist and behaviorist in nature
\footnote{An early essay by
Amartya Sen on the philosophical foundations of traditional decision theory
makes the behaviorist reading of decision theory clear
\cite{sen_behaviour_1973}.  Kaushik Basu also discusses the behaviorist
nature of decision theory {\cite[pgs. 53--54]{basu_revealed_1980}}. 
Finally, Donald Davidson appeals to decision theory to motivate an externalist
epistemology in \cite{davidson_could_1995}.}.

In {\cite{van_benthem_reflectionsepistemic_1991}}, Johan van Benthem proposed
a research program to find logics for explicit knowledge, providing the first
suggestion of an internalist perspective on epistemic logic.  Subsequently,
Sergei Artemov and Elena Nogina proposed a logic of explicit justification,
which has come to be known as \tmtextit{Justification Logic} (JL)
{\cite{artemov_introducing_2005}}, based on Artemov's \tmtextit{Logic of
Proofs} {\cite{artemov_logic_1994}}.  While the original semantics of
JL were based on interpretability into Peano Arithmetic, Melvin Fitting proposed Kripke
semantics for JL in {\cite{fitting_logic_2005}}.  Recently, Fernando
Vel\'azquez-Quesada and Johan van Benthem have developed a simpler framework
for explicit epistemics in {\cite{van_benthem_inference_2009}}, entitled the
\tmtextit{Dynamics of Awareness}.  This work was based on Joseph Halpern and
Ronald Fagin's original \tmtextit{Awareness Logic} {\cite{fagin_belief_1987}}.

@@

We adopt a arguments-on-paper-thesis perspective on epistemic logic in this
paper.  We consider a \tmtextit{good argument} to be a logical derivation
from propositions present in a knowledge base in this setting.  \
Special attention will be given to \tmtextit{sound derivations}, which will be
thought of as a form of knowledge.  In \emph{Awareness Logic}, we will interpret
awareness of a formula as membership in a knowledge base.  We prove
completeness for basic awareness logic and a hybrid logic extension.  Logics
of multiple knowledge bases are also presented: a simplified form of
Justification Logic, a logic with neighborhood semantics, an a logic with
modalities for quantifying over knowledge bases.  We conclude with an
application to naturalized epistemology found in the psychological
literature.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End:
